---
title: "Chess Game Simulation"
author: "Alex Mangiapane and Sean Stuhlsatz"
output: html_notebook
---


Here, we will simulate the results of chess games by first initializing the ratings of players, then simulating the outcomes through a function of the difference in player ratings.

```{r}
ratings <- as.vector(runif(10, min = 1400, max = 1600))
ratings = sort(ratings)
#This rating vector represents the true underlying ratings of players. However, we will assume each player starts with a provisional rating of 1500. 
ratings = ratings - (mean(ratings) - 1500)
```

Now we can define the probability function, which inputs two ratings and outputs the chance of player 1 winning.

```{r}
no_tie <- function(r1, r2, coef = -log(10)/400){
  prob <- 1/(1+exp(coef*(r1-r2)))
  return(prob)
}
```

We test this function on several sample ratings:
```{r}
no_tie(1500,1400)
no_tie(1400,1300)
no_tie(2000,1000)
```

This is the initial logistic choice of probability function, but we would like to also use a function which allows for draws. We use the logistic function suggested by Glicko, which is only a slight deviation on the initial Bradley-Terry function. 
```{r}
#glicko_prob <- function(r1, r2, s){
  #prob <- (10^((r1-r2)/400))^s/(1+10^((r1-r2)/400))
  #return(prob)
#}
#s {0, 1/2, 1}
```

Again, we test several sample ratings:

```{r}
#glicko_prob(1500,1400,1)
#glicko_prob(1500,1400,1/2)
#glicko_prob(1500,1400,0)
#Don't add to 1... How did we make sense of this?
```

Let's first try using the no_tie function to generate outcomes of games between random players: 

```{r}
rbinom(1, 1, no_tie(ratings[1], ratings[2]))
```

If this function returns a 0, that corresponds to the second player winning. If it returns a 1, that corresponds to the first player winning. This is randomly chosen as a Bernoulli trial with probability given by our no_tie function. 

Below, we create a matrix displaying the probability of a victory in any given matchup.

```{r}
p_matrix_real <- matrix(, nrow = 10, ncol = 10)
for(j in 1:10){
   for(i in 1:10){
        p_matrix_real[i,j] = no_tie(ratings[i], ratings[j])
   }
}

```

Now let us simulate the results of games where each player plays every other player, and store the results in a matrix.

```{r}
s_matrix <- matrix(, nrow = 10, ncol = 10)
for(j in 1:10){
   for(i in 1:j){
        s_matrix[i,j] = 2*rbinom(1, 1, no_tie(ratings[i], ratings[j]))-1
   }
}
for(j in 1:10){
   for(i in j:10){
        s_matrix[i,j] = -s_matrix[j,i]
   }
}
for(j in 1:10){
   for(i in j:j){
        s_matrix[i,j] = 0
   }
}
```

In the s_matrix, we store a 1 in the (i, j) entry if player i beat player j, and a -1 if player i lost to player j. We fill in 0's along the diagonal for completeness. Next, we need to update the players provisional rating of 1500 with the Elo update algorithm, which compares probability of wins to the actual outcomes. We let K = 32 be a constant updating factor. 

```{r}
priors <- c(rep(1500, 10))
sum = 1
t = 100
euclid_norm <- c(rep(NULL, t))

for(n in 1:t){
p_matrix <- matrix(, nrow = 10, ncol = 10)
for(j in 1:10){
   for(i in 1:10){
        p_matrix[i,j] = no_tie(priors[i], priors[j])
   }
}
s_matrix <- matrix(, nrow = 10, ncol = 10)
for(j in 1:10){
   for(i in 1:j){
        s_matrix[i,j] = 2*rbinom(1, 1, no_tie(ratings[i], ratings[j]))-1
   }
}
for(j in 1:10){
   for(i in j:10){
        s_matrix[i,j] = -s_matrix[j,i]
   }
}
for(j in 1:10){
   for(i in j:j){
        s_matrix[i,j] = 0
   }
}
#s_matrix[1,]
#sum(s_matrix[1,])
#Note that this value is the number is wins - losses. Since wins + losses = 9 games, we can sum these values and divide by two to get wins, and the rest are losses.
posteriors <- c(rep(NULL, 10))
for (i in 1:10){
  posteriors[i] <- priors[i] + 32/(sum)*((sum(s_matrix[i,])+9)/2 - sum(p_matrix[i,]) + 0.5)
}
priors = posteriors
euclid_norm[n] = sum((ratings - priors)^2)
sum = sum + 1

}

```


```{r}

# ratings
# mean(ratings)
# posteriors
# sum((ratings - posteriors)^2)
plot(1:t, euclid_norm)
```


```{r}
#Checking convergence
euclid_norm[t]
ratings
priors
```


The above code works for the scenario where everyone plays everyone else in a given round. Next, we will implement different matching schedules.

```{r}
GetBTProb <- function(skills, coeff=1) {
  n <- length(skills)
  skillMat <- matrix(rep(skills,n),nrow=n,byrow=TRUE)
  probMat <- as.matrix((1+exp(coeff*(skillMat-t(skillMat))))^(-1))
  return(probMat)
}
# Function for computing Thurstone-Mosteller probabilities.
GetTMProb <- function(skills, coeff=1) {
  n <- length(skills)
  probMat <- matrix(0,n,n)
  for(j in 1:n) probMat[,j] <- pnorm(skills,
                                     mean = skills[j],
                                     sd=1/coeff)
  probMat
}
```

Note that the above code generalizes the no-tie function, returning a probability matrix for each matchup.

```{r}
q = log(10)/400
p_matrix <- GetBTProb(ratings, q)
#p_matrix[1,4]
```

Now, we introduce a MatchRandomly function, which allows us to create $n$ randomly selected pairs from a group of $m$ players. 

```{r}
MatchRandomly <- function(m,n=1000) {
  matchups <- data.frame("i" = rep(0,n))
  matchups$i <- ceiling(runif(n,min=0,max=m))
  matchups$j <- ((matchups$i-1 + ceiling(runif(n,min=0,max=m-1))) %% m)+1
  return(matchups)
}
```

We test this function below:

```{r}
MatchRandomly(length(ratings), 10)
```

We can now use this function to randomly generate games by which we can simulate results and see how the ratings converge, again with the Elo update. 

```{r}
priors_mr <- c(rep(1500, 10))
euclid_norm_mr <- c(rep(NULL, t))
time_mr <- 1
matchups <- MatchRandomly(length(ratings), t)
mr_results <- matrix(, nrow = 1, ncol = t)
for(j in 1:t){
  mr_results[1,j] = rbinom(1, 1, p_matrix[matchups[j,1], matchups[j,2]])
}
for(i in 1:t){
  priors_mr[matchups[i, 1]] <- priors_mr[matchups[i, 1]] + 32/time_mr*(mr_results[1, i] - p_matrix[matchups[i,1], matchups[i,2]])
  priors_mr[matchups[i, 2]] <- priors_mr[matchups[i, 1]] + 32/time_mr*(-mr_results[1, i] + p_matrix[matchups[i,1], matchups[i,2]])
  euclid_norm_mr[i] = sum((ratings - priors_mr)^2)
  time_mr = time_mr + 1
}

```

```{r}
plot(1:t, euclid_norm_mr)
```

