---
title: "Chess Game Simulation"
author: "Alex Mangiapane and Sean Stuhlsatz"
output: html_notebook
---


Here, we will simulate the results of chess games by first initializing the ratings of players, then simulating the outcomes through a function of the difference in player ratings.

```{r}
ratings <- as.vector(runif(10, min = 1400, max = 1600))
ratings = sort(ratings)
ratings[1]
```

Now we can define the probability function, which inputs two ratings and outputs the chance of player 1 winning.

```{r}
no_tie <- function(r1, r2){
  prob <- (10^((r1-r2)/400))/(1+10^((r1-r2)/400))
  return(prob)
}
```

We test this function on several sample ratings:
```{r}
no_tie(1500,1400)
no_tie(1400,1300)
no_tie(2000,1000)
```

This is the initial logistic choice of probability function, but we would like to also use a function which allows for draws. We use the logistic function suggested by Glicko, which is only a slight deviation on the initial Bradley-Terry function. 
```{r}
#glicko_prob <- function(r1, r2, s){
  #prob <- (10^((r1-r2)/400))^s/(1+10^((r1-r2)/400))
  #return(prob)
#}
#s {0, 1/2, 1}
```

Again, we test several sample ratings:

```{r}
#glicko_prob(1500,1400,1)
#glicko_prob(1500,1400,1/2)
#glicko_prob(1500,1400,0)
#Don't add to 1... How did we make sense of this?
```

Let's first try using the no_tie function to generate outcomes of games between random players: 

```{r}
rbinom(1, 1, no_tie(ratings[1], ratings[2]))
```

If this function returns a 0, that corresponds to the second player winning. If it returns a 1, that corresponds to the first player winning. This is randomly chosen as a Bernoulli trial with probability given by our no_tie function. 

Below, we create a matrix displaying the probability of a victory in any given matchup.

```{r}
p_matrix <- matrix(, nrow = 10, ncol = 10)
for(j in 1:10){
   for(i in 1:10){
        p_matrix[i,j] = no_tie(ratings[i], ratings[j])
   }
}

```

Now let us simulate the results of games where each player plays every other player, and store the results in a matrix.

```{r}
s_matrix <- matrix(, nrow = 10, ncol = 10)
for(j in 1:10){
   for(i in 1:j){
        s_matrix[i,j] = 2*rbinom(1, 1, no_tie(ratings[i], ratings[j]))-1
   }
}
for(j in 1:10){
   for(i in j:10){
        s_matrix[i,j] = -s_matrix[j,i]
   }
}
for(j in 1:10){
   for(i in j:j){
        s_matrix[i,j] = 0
   }
}
```

In the s_matrix, we store a 1 in the (i, j) entry if player i beat player j, and a -1 if player i lost to player j. We fill in 0's along the diagonal for completeness. 