---
title: "R Notebook"
author: "Alex Mangiapane and Sean Stuhlsatz"
output: html_notebook
---

This notebook will be dedicated to an analysis of various ranking system methods on lichess.org data. First, we import the dataset:

```{r}
library(tidyverse)
lichess <- read.csv("games.csv")
```

Let us do some preliminary analysis of this data:

```{r}
summary(lichess)
head(lichess)
```

We now create a new column of data containing the rating difference between the two players. This figure will be calculated as $white rating - black rating$. 

```{r}
rating_dif <- lichess$white_rating - lichess$black_rating
lichess$rating_dif = rating_dif
```

Now, we filter out the unrated games, so that we can track how ratings update. 

```{r}
#Loops through and stores all the indeces of unrated games, then removes them from the dataframe. Stores these rated games in a new dataframe. 
dv <- c()
counter = 1
for(i in 1:nrow(lichess)){
  if(lichess$rated[i] == "FALSE" | lichess$rated[i] == "False"){
    dv[counter] <- i
    counter = counter + 1
  }
}
rated_lichess <- lichess[-dv,]
```

```{r}
#cor(rated_lichess$turns, abs(rated_lichess$rating_dif))
```

We now sort the games by the time they were created so that we can chronologically track ratings. 

```{r}
ordered_lichess <- rated_lichess[order(rated_lichess$created_at),]
```

With the data sorted, we can now apply our ranking system methods to the dataset and compare their performance in predicting outcomes. We start with TrueSkill, where we experiment with various values for $\epsilon$ to see which performs best. First, let us import the TrueSkill update function:

```{r}
psum = 0
vfuncdraw <- function(a, b){
   return((dnorm(-a-b)-dnorm(a-b))/(pnorm(a-b)-pnorm(-a-b)))
}
vfunc <- function(a, b){
   return(dnorm(a-b)/pnorm(a-b))
}
wfuncdraw <- function(a, b){
   return((vfuncdraw(a, b))^2 + ((a-b)*(dnorm(a-b)) + (a+b)*(dnorm(a+b)))/(pnorm(a-b)-pnorm(-a-b)))
}
wfunc <- function(a, b){
   return(vfunc(a, b)*(vfunc(a, b) + a - b))
}
RateTrueskill <- function(games, p_data, eps = 0, beta = 300){
   for(n in 1:nrow(games)){
     
p1 = 1-pnorm(eps, p_data[games[n, 1], 1]-p_data[games[n, 2], 1], p_data[games[n, 1], 2]+p_data[games[n, 2], 2])
   #player 2 wins probability
p2 = 1-pnorm(eps, -p_data[games[n, 1], 1]+p_data[games[n, 2], 1], p_data[games[n, 1], 2]+p_data[games[n, 2], 2])
   #draw probability
p3 = max(1 - p1 - p2, 0)
     
      c = sqrt(2*beta^2 + p_data[games[n, 1], 2]^2 + p_data[games[n, 2], 2]^2)
      
      if(games[n, 3] == 0.5){
        #Check this part
         p_data[games[n, 1], 1] = p_data[games[n, 1], 1] + ((p_data[games[n, 1], 2]^2)/c)*vfuncdraw(((p_data[games[n, 1], 1]) - (p_data[games[n, 2], 1]))/c, eps/c)
      
      p_data[games[n, 2], 1] = p_data[games[n, 2], 1] - ((p_data[games[n, 2], 2]^2)/c)*vfuncdraw(((p_data[games[n, 1], 1]) - (p_data[games[n, 2], 1]))/c, eps/c)
      
      p_data[games[n, 1], 2] = p_data[games[n, 1], 2]*(1-(((p_data[games[n, 1], 2])^2)/(c^2))*wfuncdraw((p_data[games[n, 1], 1] - p_data[games[n, 2], 1])/c, eps/c))
      
      p_data[games[n, 2], 2] = p_data[games[n, 2], 2]*(1-(((p_data[games[n, 2], 2])^2)/(c^2))*wfuncdraw((p_data[games[n, 1], 1] - p_data[games[n, 2], 1])/c, eps/c))
      
      psum = psum + log(p3)
      print(p3)
      #print(psum)
      }else{
           p_data[games[n, 1], 1] = p_data[games[n, 1], 1] + ((p_data[games[n, 1], 2]^2)/c)*vfunc(((p_data[games[n, 1], 1]) - (p_data[games[n, 2], 1]))/c, eps/c)
      
      p_data[games[n, 2], 1] = p_data[games[n, 2], 1] - ((p_data[games[n, 2], 2]^2)/c)*vfunc(((p_data[games[n, 1], 1]) - (p_data[games[n, 2], 1]))/c, eps/c)
      
      p_data[games[n, 1], 2] = p_data[games[n, 1], 2]*(1-(((p_data[games[n, 1], 2])^2)/(c^2))*wfunc((p_data[games[n, 1], 1] - p_data[games[n, 2], 1])/c, eps/c))
      
      p_data[games[n, 2], 2] = p_data[games[n, 2], 2]*(1-(((p_data[games[n, 2], 2])^2)/(c^2))*wfunc((p_data[games[n, 1], 1] - p_data[games[n, 2], 1])/c, eps/c))
      
      psum = psum + log(p1)
      #print(psum)
      print(p1)
      }
   }
    #print(psum)
   return(p_data)
}
```

Now, we loop through the games.

```{r}
#Initializing an empty dataframe to hold players information
usernames <- c()
ratings <- c()
variances <- c()


#Now we loop through games; any time we encounter a new user, we assign them a rating of 1500 and variance of 300. If they then play multiple games, we update these parameters with the trueskill update.

ordered_lichess <- ordered_lichess %>% select(white_id, black_id, rating_dif, winner)

#NEED A THREE COLUMN DATAFRAME, WHERE THE FIRST COLUMN HAS THE INDEX IN USERNAMES OF THE WINNER, SECOND HAS THE USERNAME INDEX OF THE LOSER, AND THE THIRD HAS THE RESULT (EITHER 1/2 OR 1)

for(n in 1:nrow(ordered_lichess)){
  if(ordered_lichess$white_id[n] %in% usernames == FALSE){
    usernames <- append(usernames, ordered_lichess$white_id[n])
    ratings <- append(ratings, 1500)
    variances <- append(variances, 300)
  }
  if(ordered_lichess$black_id[n] %in% usernames == FALSE){
    usernames <- append(usernames, ordered_lichess$black_id[n])
    ratings <- append(ratings, 1500)
    variances <- append(variances, 300)
  }
}

player_data <- data.frame(ratings, variances, usernames)

index_one <- c()
index_two <- c()
result <- c()
for(n in 1:nrow(ordered_lichess)){
  index_one <- append(index_one, match(ordered_lichess[n, 1], usernames))
  index_two <- append(index_two, match(ordered_lichess[n, 2], usernames))
  if(ordered_lichess$winner[n] == "draw"){
    result <- append(result, 0.5)
  }else{
    if(ordered_lichess$winner[n] == "white"){
      result <- append(result, 1)
    }else{
      result <- append(result, 0)
    }
  }
}

#Now we have to manipulate the dataframe to have 

```


```{r}
#Now, we create the dataframe expected by the RateTrueskill function by cbinding these vectors. 

game_data <- data.frame(index_one, index_two, result)
for(i in 1:nrow(game_data)){
  if(game_data$result[i] == 0){
    game_data$result[i] = 1
    dummy <- game_data[i, 1]
    game_data[i, 1] = game_data[i, 2]
    game_data[i, 2] = dummy
  }
}
```

Now, the data is in the correct form, so we can apply the RateTrueskill function. 
```{r}
final_ratings_zero <- RateTrueskill(game_data, player_data, 0, 300)
```

#Why are some of the ratings saying NaN? Is there a problem when there is a draw? Check on this...

Let us try this same calculation with a different value of epsilon:

```{r}
final_ratings_5 <- RateTrueskill(game_data, player_data, 5, 300)
```


