library(tictoc)

# This .R file is laid out as follows: the first set of 
# functions take in a vector of skills and output a matrix
# of pairwise win-loss probabilities; the second set of 
# functions take in a number of players (and sometimes
# other parameters) and output a schedule of games (in the
# form of a dataframe with columns for which player is player
# i and which is player j and a row for each game); then the
# function MakeGames takes in a matrix of pairwise win-loss 
# probabilities and a schedule of games and outputs a 
# dataframe containing the results.


###################################
###   Functions for computing   ###
###         probabilities       ###
###################################

# Old method for computing Bradley-Terry probabilities.
# Less vectorized (ergo probably less efficient) than
# GetBTProb.
```{r}
GetBTProbScalar <- function(skills,coeff=1) {
  n <- length(skills)
  probMat <- matrix(0,n,n)
  for(i in 1:n) {
    for(j in 1:n) {
      probMat[i,j] <- 1/(1+exp(-coeff*(skills[i]-skills[j]))) 
    }
  }
  return(probMat)
}
```


# New method for computing Bradley-Terry probabilities.
# To get Glickman's Bradley-Terry model, set coeff=ln10/400.
```{r}
GetBTProb <- function(skills, coeff=1) {
  n <- length(skills)
  skillMat <- matrix(rep(skills,n),nrow=n,byrow=TRUE)
  probMat <- (1+exp(coeff*(skillMat-t(skillMat))))^(-1)
  return(probMat)
}
# Function for computing Thurstone-Mosteller probabilities.
GetTMProb <- function(skills, coeff=1) {
  n <- length(skills)
  probMat <- matrix(0,n,n)
  for(j in 1:n) probMat[,j] <- pnorm(skills,
                                     mean = skills[j],
                                     sd=1/coeff)
  probMat
}
```


```{r}
GetBTProb(skills, log(10)/400)
skills
```

####################################
###   Functions for scheduling   ###
###            games             ###
####################################


#Chooses a player for each game unif. at random, then
#chooses another play for them to play against unif.
#at random.
```{r}
MatchRandomly <- function(m,n=1000) {
  matchups <- data.frame("i" = rep(0,n))
  matchups$i <- ceiling(runif(n,min=0,max=m))
  matchups$j <- ((matchups$i-1 + ceiling(runif(n,min=0,max=m-1))) %% m)+1
  return(matchups)
}
```

#Assigns players to matches an opponents randomly, but
#each player plays exactly one game in the first m/2.
#If m is odd, a randomly-selected player plays 2 games.
#Param n is number of rounds, not number of games
```{r}
MatchAllPlayers <- function(m,n=1) {
  c <- ceiling(m/2)
  f <- floor(m/2)
  matchups <- data.frame("i" = rep(0,c*n))
  iTeams <- rep(0,c)
  jTeams <- rep(0,c)
  for(k in 0:(n-1)) {
    iTeams <- sample(1:m,c,replace=FALSE)
    jTeams[1:f] <- sample((1:m)[-iTeams],f,replace=FALSE)
    if(f<c) jTeams[c] <- sample((1:m)[-iTeams[c]],1,replace=FALSE)
    matchups$i[(k*c+1):(k*c+c)] <- iTeams
    matchups$j[(k*c+1):(k*c+c)] <- jTeams
  }
  return(matchups)
}
```



# Every possible pair of players, n times.
# If randomize=TRUE, return in a random order (within
# rounds).
# I suspect there's a one-line way to do this, but
# I couldn't find it.
```{r}
MatchAllPairs <- function(m,n=1,randomize=FALSE) {
  r <- m*(m-1)/2
  matchups <- data.frame("i"=c(),"j"=c())
  for(k in 1:n) {
    for(l in 1:(m-1)) {
      matchups <- rbind(matchups,
            data.frame("i"=rep(l,m-l),
                       "j"=(l+1):m))
    }
    if(randomize) matchups[(1+((k-1)*r)):(k*r),] <- 
        (matchups[(1+((k-1)*r)):(k*r),])[sample(1:r,r),]
  }
  print(r)
  return(matchups)
}
```


# UNFINISHED
# Schedules players based on skills. If closest=TRUE,
# tries to schedule players with similar skills; o.w.
# tries to schedule players with dissimilar skills.
# For games under competitive matchmaking, can use
# current skill estimates for the skills argument
```{r}
MatchBySkill <- function(skills,closest=TRUE,n) {
  temp <- data.frame("i"=c(),
                     "j"=c())
  for(k in 1:n) {
    temp <- data.frame("i"=c(),
                       "j"=c())
    while(length(temp$i) < floor(m/2)) {
      options <- (1:m)[-c(temp$i,temp$j)]
      temp$i[length(temp$i)+1] <- options[which.max(skills[options])]
      options <- (1:m)[-c(temp$i,temp$j)]
      temp$j[length(temp$j)+1] <- options[which.max(skills[options])]
    }
    #UNFINISHED
  }
}
```

####################################
###   Functions for simulating   ###
###            games             ###
####################################


# Simulates the games scheduled in dataframe games between
# players with pairwise win/loss probabilities taken from
# matrix prob.
#
# I tried to vectorize this function, but performance
# got worse. (In fact, the performance of this function
# was very strange -- I'm at a loss for why the code I
# have ran faster than other code I tried.)

```{r}
MakeGames <- function(prob,games) {
  n <- length(games$i)
  
  #tic("Assign probabilities")
  games$p <- rep(0,length(games$i))
  for(k in 1:n) games$p[k] <- prob[games$i[k],games$j[k]]
  #toc()
  
  #tic("Play the game")
  for(k in 1:n) {
    games$results[k] <- rbinom(1,1,prob[games$i[k],games$j[k]])
  }
  #toc()
  
  return(games)
}
```

# Makes n tournaments between m players with matrix of
# win-loss probabilities prob. Each tournament works as
# follows: players are randomly assigned to games; the
# winners of one round progress to the next.

```{r}
MakeTournament <- function(prob,n) {
  #UNFINISHED
}
```

###################################
###    Tests for functions      ###
###################################

```{r}
skills <- runif(10,min=1500,max=2500)
#skills2 <- c(1,1,2,3,4,4,0)
q=log(10)/400
```

### Probability calculation tests

# Confirm that the vectorized BTProb function produces the
# same output as the non-vectorized one
```{r}
GetBTProb(skills,q)-GetBTProbScalar(skills,q)

# Test functions


GetTMProb(c(0,1,1.5,1.96))
GetBTProb(c(0,1,1.5,1.96))

GetTMProb(skills,q)


### Scheduling tests
MatchRandomly(10,20)
MatchRandomly(5,12)
MatchRandomly(4)

MatchAllPlayers(6,5)
MatchAllPlayers(5,3)
MatchAllPlayers(10)

MatchAllPairs(4,2)
MatchAllPairs(4,2,randomize=TRUE)
MatchAllPairs(5)
```


### Game-making tests

```{r}
MakeGames(prob=GetBTProb(skills,q),
          games = MatchRandomly(length(skills),25))

MakeGames(prob=GetTMProb(skills,q),
          games = MatchAllPairs(length(skills),1))
MakeGames(prob=GetTMProb(skills,q),
          games = MatchAllPairs(length(skills),1))

games <- MakeGames(prob = GetBTProb(skills,q),
                   games = MatchRandomly(length(skills),10000))
sum(games$p*games$results)/sum(games$results)
sum((1-games$p)*games$results)/sum(1-games$results)
for(i in 0:19) {
  print(mean(games$results[which((games$p>i*0.05) & (games$p < (i+1)*0.05))]))
}

MakeGames(GetBTProb(skills2),MatchRandomly(length(skills2),100))
```




#Benchmarking Bradley-Terry prob functions
```{r}
x <- matrix(0,nrow=4000,ncol=4000)

tic()
x <- GetBTProb(1:4000,q)
toc()

tic()
x <- GetBTProbScalar(1:4000,q)
toc()

```


#Mild difference observed. Can't be bothered to confirm
#at this juncture through more rigorous benchmarking





TO DO!!!

Work out the update function using Sean's scheduling methods, particularly after each game. Pick one rating system (Elo, Glicko, Trueskill....) and try to optimize both the initialization of the K-factor and its subsequent decay. Graph the euclidean norm between the actual ratings of the players over time (which should hopefully decay as something like f(x) = 1/x). We must then compare rating systems against each scheduling method one at a time, and perhaps find that different ratings systems perform differently on different methods of scheduling. 
Make it look pretty as well... format with comments explaining what each chunk of code is doing, in a similar way to the way you worked through the earlier "no-tie" function part.

```{r}
MatchRandomly(10, 1)
```


```{r}
#Alex functions that are useful
no_tie <- function(r1, r2, coef = -q){
  prob <- 1/(1+exp(coef*(r1-r2)))
  return(prob)
}
p_matrix_elo <- matrix(, nrow = length(skills), ncol = length(skills))
for(j in 1:length(skills)){
   for(i in 1:length(skills)){
        p_matrix_elo[i,j] = no_tie(skills[i], skills[j])
   }
}
 #p_matrix_elo
```


```{r}
#Elo probability function and update
lower_bound <- 1500
upper_bound <- 2500
probs <- GetBTProb(skills)
skills <- c(seq(1500, 2400, 100))
skills = skills-mean(skills)+(lower_bound+upper_bound)/2
```

Here, we will implement the update process by using the MatchRandomly function. Note that our estimate of players true ratings is stored in the ratings variable, so we can only use this variable to determine probabilities, as well as in our update equation. 


```{r}
ratings <- c(rep((upper_bound+lower_bound)/2, length(skills)))
time = 1
t = 1000
euclid_norm <- c(rep(NULL, t))
for(i in 1:t){
game <- MatchRandomly(length(skills), 1)
x <- as.integer(game[1])
y <- as.integer(game[2])
r <- rbinom(1, 1, p_matrix_elo[x, y])


ratings[x] <- ratings[x] + 5*(r - p_matrix_elo[x,y])
ratings[y] <- ratings[y] + 5*(p_matrix_elo[x,y] - r)
euclid_norm[time] = sum((ratings - skills)^2)
time = time + 1
}

```


```{r}
ratings 
mean(ratings)
skills

plot(1:t, euclid_norm)
```

