---
title: "Chess Game Simulation"
author: "Alex Mangiapane and Sean Stuhlsatz"
output: html_notebook
---


This document is meant to encode the TrueSkill update formula so that we can simulate games and examine convergence rates between this method and others, such as Elo and Glicko. 

```{r}
rating_means <- sort(as.vector(runif(10, min = 1400, max = 1600)))
rating_means = rating_means - (mean(rating_means) - 1500)
rating_vars <- c(rep(300, 10))
player_data <- data.frame(rating_means, rating_vars)
priors <- data.frame(c(rep(1500, 10)), c(rep(300, 10)))
```

```{r}
player_data
```

```{r}
#No win or loss
vfunc <- function(a, b){
   return(dnorm(a-b)/pnorm(a-b))
}
#No draws allowed w function
wfunc <- function(a, b){
   return(vfunc(a, b)*(vfunc(a, b) + a - b))
}
RateTrueskill <- function(games, p_data, eps, beta){
   for(n in 1:nrow(games)){
      c = sqrt(2*beta^2 + p_data[games[n, 1], 2]^2 + p_data[games[n, 2], 2]^2)
      
      p_data[games[n, 1], 1] = p_data[games[n, 1], 1] + ((p_data[games[n, 1], 2]^2)/c)*vfunc(((p_data[games[n, 1], 1]) - (p_data[games[n, 2], 1]))/c, eps/c)
      
      p_data[games[n, 2], 1] = p_data[games[n, 2], 1] - ((p_data[games[n, 2], 2]^2)/c)*vfunc(((p_data[games[n, 1], 1]) - (p_data[games[n, 2], 1]))/c, eps/c)
      
      p_data[games[n, 1], 2] = p_data[games[n, 1], 2]*(1-(((p_data[games[n, 1], 2])^2)/(c^2))*wfunc((p_data[games[n, 1], 1] - p_data[games[n, 2], 1])/c, eps/c))
      
      p_data[games[n, 2], 2] = p_data[games[n, 2], 2]*(1-(((p_data[games[n, 2], 2])^2)/(c^2))*wfunc((p_data[games[n, 1], 1] - p_data[games[n, 2], 1])/c, eps/c))
   }
   return(p_data)
}
```

Let's test out this update function on two artificial games:

```{r}
# winners <- c(1, 3)
# losers <- c(2, 4)
# df <- data.frame(winners, losers)
#df
#priors <- RateTrueskill(df, priors, 0, 300)
```

Now, we can implement the match randomly function to simulate games and test how well the ratings converge to the rating means. 

```{r}
MatchRandomly <- function(m,n=1000) {
  matchups <- data.frame("i" = rep(0,n))
  matchups$i <- ceiling(runif(n,min=0,max=m))
  matchups$j <- ((matchups$i-1 + ceiling(runif(n,min=0,max=m-1))) %% m)+1
  return(matchups)
}
#m is the number of playes, n is the number of games
```

We will first create 1000 simulated matchups:

```{r}
games <- MatchRandomly(length(rating_means))
```

Next, we must go in and simulate results for each of these games. Note that determining a winner using Trueskill is more complicated than Elo because of the inclusion of a draw parameter $\epsilon$. Should we wish to exclude the possibility for draws, we simply set $\epsilon = 0$. In this simulation, we will create a vector with 1's if the first player in the games variable wins, and 0's if the second player wins. 0.5 if drawn.  

```{r}
eps = 0
   results <- c()
for(n in 1:nrow(games)){
   #player 1 wins probability
   p1 = 1-pnorm(eps, rating_means[games[n, 1]]-rating_means[games[n, 2]], rating_vars[games[n, 1]]+rating_vars[games[n, 2]])
   #player 2 wins probability
   p2 = 1-pnorm(eps, rating_means[games[n, 2]]-rating_means[games[n, 1]], rating_vars[games[n, 1]]+rating_vars[games[n, 2]])
   #draw probability
   p3 = 1 - p1 - p2
   
   if(rbinom(1, 1, p3) == 1){
      results = append(results, 0.5)
   } else {
      results = append(results, rbinom(1, 1, p1/(p1+p2)))
      }
}
#games$results = results
results


```

```{r}
#create_results(games)
# test1 = 0.6
# test2 = 0.3
# test3 = 0.1
# 
# newv = c()
# 
# for(i in 1:10){
#    if(rbinom(1, 1, test3) == 1){
#       newv = append(newv, 0.5)
#    } else{
#       newv = append(newv, rbinom(1, 1, test1/(test1 + test2)))
#    }
# }
# print(newv)
```


